output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
# Add any other packages you need to load here.
library(tidyverse)      # data wrangling and ggplot2
library(rsample)        # initial_split, training, testing 
library(yardstick)      # rmse, mae, accuracy, conf_mat 
library(broom)          # tidy, glance, augment
library(rpart)          # regression trees, printcp, plotcp 
library(randomForest)   # random forest, importance
library(nnet)           # neural network, multinom
library(discrim)        # LDA, KDA
library(GGally)
# Read in the data
property.train <- read_csv("https://www.massey.ac.nz/~jcmarsha/161777/assessment/data/property_train_val.csv")
property.test <- read_csv("https://www.massey.ac.nz/~jcmarsha/161777/assessment/data/property_test.csv")
wheat.train <- read_csv("https://www.massey.ac.nz/~jcmarsha/161777/assessment/data/wheat_train_val.csv") |>
  mutate(variety = factor(variety))
wheat.test <- read_csv("https://www.massey.ac.nz/~jcmarsha/161777/assessment/data/wheat_test.csv")
```

Replace `YOUR NAME HERE` in the author field before knitting. Save your
final HTML report as `report_studentid.html`.

## Exercise 1: Property Value Prediction

### Step 1 — Exploratory Data Analysis

```{r}
# 1a. Check for missing values
# confirms 100% present, no missingness
naniar::vis_miss(property.train)     

# 1b. Summary statistics for all variables
# 310 observations, 6 predictors and 1 target
# The maximum value of transit_distance is 6488, while the median value is only 492, indicating a large disparity; 
skimr::skim(property.train)         

# 1c. Faceted scatterplots: price_per_unit vs every predictor
# In almost all of the plots, there is a  price_per_unit outlier(117.5). 
# Latitude and longitude -spatial clustering patterns. nearby_stores-positive , property_age-negative
# a log transformation will be applied to transit_distance to examine whether a clearer relationship pattern can be observed.
property.train |>
  pivot_longer(-c(price_per_unit),
               names_to  = "name",
               values_to = "value") |>
  ggplot() +
  geom_point(mapping=aes(x = value, y = price_per_unit), alpha = 0.5) +
  facet_wrap(vars(name), scales = "free_x") +    
  labs(x = NULL, y = "price_per_unit",
       title = "price_per_unit vs each predictor (original scale)")


# 1d. Transit_distance: original vs log scale
# Result: log scale gives a more linear relationship — log transform appropriate
ggplot(property.train) +
  geom_point(mapping=aes(x = transit_distance, y = price_per_unit),
             col = "purple", alpha = 0.5) +
  labs(title = "Original Scale")

ggplot(property.train) +
  geom_point(mapping=aes(x = transit_distance + 1, y = price_per_unit),
             col = "blue", alpha = 0.5) +
  scale_x_log10() +
  labs(title = "Log Scale (transit_distance + 1)")

```

### Step 2 — Data processing

```{r}
# train data split(analysis and validation)
# We use 75% (≈232 rows) for fitting and 25% (≈78 rows) for RMSE evaluation. we do NOT touch property.test until the very final prediction step.

set.seed(1234)
prop_split    <- initial_split(property.train, prop = 0.75)
prop_analysis <- training(prop_split)  
prop_val      <- testing(prop_split)    
```

### Step 3 — Modelling Techniques Tried

#### 3.1 Model A — Linear Regression

##### Round 1: Fit linear regression with all predictors

```{r}
# Train a model on prop_analysis, and then predict prop_val  and evaluate the RMSe
prop.lm <- linear_reg()|> fit(price_per_unit ~., data = prop_analysis)

# The R² value is 0.55, indicating that the model explains 55% of the variation in the target variable .
glance(prop.lm)                    

# Select p.value > 0.05 (likely candidates for removal)
tidy(prop.lm) |> filter(p.value > 0.05)

# Calculate RMSE
prop.lm.pred<-prop.lm  |> augment(prop_val )
prop.lm.pred |>  rmse(truth = price_per_unit, estimate = .pred)

```

##### Round 2: Establish a Baseline Model

```{r}
# Fit a null model (intercept only) to represent the average price
prop.base <- linear_reg() |> 
  fit(price_per_unit ~ 1, data = prop_analysis)

# Check the model summary to confirm the intercept 
prop.base |> tidy()

# Evaluate Baseline Performance
prop_val |>
  mutate(mean_price = 37.54353)|>
  rmse(truth = price_per_unit, estimate = mean_price)

```

##### Round 3: Reduced model(-longitude)

```{r}
# Based on the tidy process in Round 1, we attempted to remove the longitude variable and then observed the RMSE value.
prop.lm2 <- linear_reg()|> fit(price_per_unit ~.-longitude, data = prop_analysis)

# The R² value is 0.55, .
glance(prop.lm2)                    

#After removing longitude, the RMSE decreased from 11.39 to 11.31.
prop.lm2.pred<-prop.lm2  |> augment(prop_val )
prop.lm2.pred |>  rmse(truth = price_per_unit, estimate = .pred)
```

##### Round 4: AIC backward stepwise selection

```{r}
# The RMSE value is the same as in Round 3,The step() function, during the automated selection process, found that 
# removing any other variables would increase the AIC value, so it ultimately chose to remove only longitude.
prop.lm <- lm(price_per_unit ~ ., data = prop_analysis)
prop.lm3 <- step(prop.lm)  

tidy(prop.lm3)    
glance(prop.lm3) 

prop.lm3.pred <- augment(prop.lm3, newdata = prop_val)
prop.lm3.pred |>  rmse(truth = price_per_unit, estimate = .fitted)


```

#### 3.2 Model B — Regression Tree

##### Round 1: Fit complex tree with small cp to see full structure

```{r}
# cp (complexity parameter):0.001
tree_complex <- rpart(price_per_unit ~ .,
                      data = prop_analysis, cp = 0.001)

# printcp() shows cross-validation error (xerror) for each cp level
# As size increases from 1 to 2, the error drops rapidly from about 1.0 to around 0.55, indicating that adding 
# a small number of branches significantly improves the model’s performance. After size > 2, the curve gradually flattens out.
cp_table <- printcp(tree_complex)
plotcp(tree_complex)  

# The smallest nsplit value (only 2 splits). Therefore, the recommended optimal Cp value is 0.0356.
cp_table |> 
  as_tibble() |> 
  mutate(limit = min(xerror + xstd)) |> 
  filter(xerror <= limit)
```

##### Round 2: Fit three trees at different cp values and compare

```{r}
# Testing cp = 0.1 (simple),0.01 (default), and the 1-SE optimal cp

tree_simple <- decision_tree(mode = "regression", cost_complexity = 0.1) |>
  fit(price_per_unit ~ ., data = prop_analysis)
tree_default <- decision_tree(mode = "regression", cost_complexity = 0.01) |>
  fit(price_per_unit ~ ., data = prop_analysis)
tree_optimal <- decision_tree(mode = "regression", cost_complexity = 0.0356) |>
  fit(price_per_unit ~ ., data = prop_analysis)
```

##### Round 3: Compare RMSE on VALIDATION SET for all three trees

```{r}
# The Default Tree performs the best, with an RMSE of 10.89, which is clearly better than the tuned Optimal Tree (11.20) and the most basic Simple Tree (11.86).
tree_simple.pred <- augment(tree_simple, new_data = prop_val)
tree_simple.pred |>  rmse(truth = price_per_unit, estimate = .pred)

tree_default.pred <- augment(tree_default, new_data = prop_val)
tree_default.pred |>  rmse(truth = price_per_unit, estimate = .pred)

tree_optimal.pred <- augment(tree_optimal, new_data = prop_val)
tree_optimal.pred |>  rmse(truth = price_per_unit, estimate = .pred)


```

#### 3.3 Model C — Random Forest

##### Round 1: Default random forest

```{r}
# Round 1: Default random forest 
# mtry defaults to p/3 = 6/3 = 2; nodesize defaults to 5 for regression
set.seed(1234)
prop.rf <- rand_forest(mode = "regression", engine = "randomForest",mtry=2,min_n=5) |>
  fit(price_per_unit ~ .,data = prop_analysis)

prop.rf.pred <- prop.rf |> augment(new_data = prop_val)
prop.rf.pred |> rmse(truth = price_per_unit, estimate = .pred)

```

##### Round 2: Random forest- change nodesize

```{r}
# After reducing min_n from 5 to 3, the RMSE on the validation set decreased from 10.398 to 10.329.
set.seed(1234)
prop.rf2 <- rand_forest(mode = "regression", engine = "randomForest",mtry=2,min_n=3) |>
  fit(price_per_unit ~ .,data = prop_analysis)

prop.rf2.pred <- prop.rf2 |> augment(new_data = prop_val)
prop.rf2.pred |> rmse(truth = price_per_unit, estimate = .pred)

```

##### Round 3: Random forest- change mtry

```{r}
# After increasing mtry from 2 to 3 (while keeping min_n = 3), the RMSE further decreased from 10.329 to 10.215.
set.seed(1234)
prop.rf3 <- rand_forest(mode = "regression", engine = "randomForest",mtry=3,min_n=3) |>
  fit(price_per_unit ~ .,data = prop_analysis)

prop.rf3.pred <- prop.rf3 |> augment(new_data = prop_val)
prop.rf3.pred |> rmse(truth = price_per_unit, estimate = .pred)

```

#### 3.4 Model D — Neural Network (Scaled Inputs Required)

```{r}
library(tidymodels)
set.seed(1234)
# This follows the recipe workflow.
# Create our recipe
prop_rec <- recipe(price_per_unit ~ .,
                   data = prop_analysis) |>
  step_normalize(all_numeric_predictors()) |>  
  prep(training = prop_analysis)               

# Create a model
spec_nn <- mlp(mode         = "regression",
               engine       = "nnet",
               hidden_units = 3,       
               penalty      = 0.01,  
               epochs       = 1000) 

# Train a model
fit_nn <- spec_nn |>
  fit(price_per_unit ~ .,  data = bake(prop_rec, prop_analysis))

# Check we're reaching converge
# convergence: 0 = converged, 1 = did not converge
fit_nn |> extract_fit_engine() |> pluck("convergence")

# Apply to test set and evaluate
pred_nn <- prop_val |> 
  bind_cols(
    predict(fit_nn, new_data = bake(prop_rec, prop_val))
  )

# Alternatively with argument
pred_nn <- fit_nn |> augment(new_data = bake(prop_rec, prop_val))

# Evaluate the model
pred_nn |> rmse(truth = price_per_unit, estimate = .pred)
```

### Step 4 — Refit on Full Training Data and Predict Test Set

```{r}
# Now that we have selected the best model using prop_analysis + prop_val, 
# we refit it on ALL 310 rows of property.train to maximize the information used before predicting. 
set.seed(1234)
prop.fina <- rand_forest(mode = "regression", engine = "randomForest", mtry = 3, min_n = 3) |> 
  fit(price_per_unit ~ ., data = property.train)

prop.final.pred<- prop.fina |> augment(new_data = property.train) |>select(-price_per_unit)

# Distribution of training outcomes
ggplot(property.train) +
  geom_density(mapping=aes(x = price_per_unit)) +
  labs(title = "Distribution of training price_per_unit")

# Distribution of predictions
# The predicted and actual values are concentrated between 20 and 60, with the distribution flattening beyond 90.
ggplot(prop.final.pred) +
  geom_density( mapping=aes(x = .pred)) +
  labs(title = "Distribution of predictions")

# histogram comparison
# The distribution comparison plot shows that the model performs consistently.
bind_rows(
  list(
    predict = prop.final.pred |> rename(price_per_unit = .pred),
    train   = property.train
  ),
  .id = "data"
) |>
  ggplot(mapping=aes(x = price_per_unit, fill = data)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 30) +
  facet_wrap(vars(data), ncol = 1, scales = "free_y") +
  labs(title = "Predictions vs Training distribution")

# It can be seen that the mean of the residuals is close to zero, indicating that the model is generally appropriate.
prop.final.pred|>skimr::skim()

# Generate predictions for property_test.csv
ex1 <- property.test |> 
  bind_cols(predict(prop.fina, new_data = property.test))

# Verify submission format: test rows + exactly one column named .pred
# check predictions are in reasonable range
ex1 |> slice_head(n = 6)  

# Write submission file
ex1 |> write_csv("ex1_studentid.csv")
```

## Exercise 1 Methodology

**Exploratory Analysis.** The training dataset contains`no missing`
values (1a). It includes 310 observations and 7 variables. Among them,
transit_distance shows a wide range, from 23 to 6,488 (mean = 1,086),
while price_per_unit ranges from 7.6 to 117.5 (mean = 37.9) (1b).

Faceted scatter plots reveal a clear negative relationship between
transit_distance and price_per_unit. In contrast, latitude, longitude,
and nearby_stores exhibit moderate positive correlations with the target
variable, whereas sale_time shows almost no linear association with the
outcome (1c).

After applying a `log transformation` to transit_distance, its
relationship with price_per_unit becomes clearer (1d).

**Data Processing.** To ensure prediction accuracy and select the
optimal model, the property.train set is split into two parts:
approximately 75% (232 observations) is used for model fitting, and the
remaining 25% (78 observations) is reserved for RMSE evaluation. The
property.test set is used solely for the final prediction step.

**Modelling Techniques Tried.** We evaluated four different types of
models, with all RMSE values computed on the validation set:

##### Model A – Linear Regression.

A full-variable `linear regression` model achieved an RMSE of `11.39`.
In this step, based on p-values, longitude was identified as the least
influential variable (Round 1).

In Round 2, a `baseline`model using the mean value as prediction was
established (`RMSE = 16.31`).

In Round 3, after `removing longitude`, the RMSE decreased slightly from
11.39 to `11.31`.

In Round 4, an `AIC-based backward stepwise` model was applied, which
again resulted in an RMSE of `11.31`. This indicates that removing
longitude is sufficient, while eliminating other variables does not
further improve performance.

##### Model B – Regression Tree.

A `complex tree (cp = 0.001)` was pruned using the 1-SE rule, resulting
in an `optimal cp = 0.0356` (2 splits) with an RMSE of `11.20`. However,
the `default tree` (cp = 0.01) performed best overall, achieving an RMSE
of `10.89`, outperforming both the pruned tree and the `simple tree` (cp
= 0.1, `RMSE = 11.86`).

##### Model C – Random Forest.

The `default configuration (mtry = 2, nodesize = 5)` achieved an RMSE of
`10.39`. Increasing `mtry to 3` (nodesize = 5) further improved
performance (`RMSE = 10.33`). Finally, adjusting both
`mtry = 3 and nodesize = 3` produced the best result with an
`RMSE of 10.215`, making it the optimal configuration.

##### Model D – Neural Network.

A single-hidden-layer `neural network` (3 hidden units, penalty = 0.01)
achieved an RMSE of `10.61`.

**Final Model Selected.** The final model selected is the
`Random Forest` with `mtry = 3 and nodesize = 3`, which achieved the
lowest `RMSE of 10.215`. We applied this model to the full dataset of
310 observations for prediction. From the comparison plots, it is
evident that the predicted distribution closely follows the actual
distribution, with a high degree of similarity in overall patterns and
the mean residuals is small(-0.09). This indicates that the model has
strong predictive performance and good practical applicability.

## Exercise 2: Wheat Variety Classification

### STEP 1: Exploratory Data Analysis

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# # 1a. Check for missing values — confirms 100% present, no missingness
naniar::vis_miss(wheat.train)  

# # 1b. Preview Training Data
# This helps inspect the predictor structure and the target variable
wheat.train |> slice_head(n = 6)

# 1c.Explore: Class Balance
# counts how many training observations belong to each variety.
# The sample sizes for all three categories (Canadian, Kama, and Rosa) are exactly the same-52
wheat.train |> count(variety)


# 1d.Summary statistics for all variables: mean, sd, quartiles, histogram 
# In each category, the maximum value differs substantially from the mean, suggesting the presence of outliers. 
# The histograms also exhibit similar distribution patterns, which will be illustrated further through visualizations.
skimr::skim(wheat.train)  

# 1e.Density Plot: distribution of each predictor coloured by variety
# The uneven data distribution makes the differences among variables across the different varieties less apparent,
wheat.train |>
  pivot_longer(-variety) |>
  ggplot() +
  geom_density(aes(x = value, col = variety)) +
  facet_wrap(vars(name), scales = "free") +
  labs(title = "Predictor distributions by wheat variety")

# 1f. Boxplots: confirm group separation in each predictor
wheat.train |>
  pivot_longer(-variety) |>
  ggplot() +
  geom_boxplot(aes(x = variety, y = value, fill = variety)) +
  facet_wrap(vars(name), scales = "free") +
  labs(title = "Boxplots of predictors by variety")

# 1g.Pairs plot: pairwise scatter coloured by variety
# Observation: many predictors are highly correlated (Most r > 0.97)
ggpairs(wheat.train, mapping = aes(col = variety, alpha = 0.4))

# 1h. Group means by variety 
wheat.train |>
  group_by(variety) |>
  summarise(across(where(is.numeric), mean)) |>
  print()

# 1i. filter outliers and then plot boxplots and Density
wheat.train.clean <- wheat.train |>
  filter(
    groove_length < 20,
    kernel_area < 50,
    kernel_asymmetry < 20,
    kernel_compactness < 2,
    kernel_length < 20,
    kernel_perimeter < 50,
    kernel_width < 10
  )
wheat.train.clean|>
  pivot_longer(-variety) |>
  ggplot() +
  geom_density(aes(x = value, col = variety)) +
  facet_wrap(vars(name), scales = "free") +
  labs(title = "Predictor distributions by wheat variety-clean")
wheat.train.clean |>
  pivot_longer(-variety) |>
  ggplot() +
  geom_boxplot(aes(x = variety, y = value, fill = variety)) +
  facet_wrap(vars(name), scales = "free") +
  labs(title = "Boxplots of predictors by variety-clean")


```

### Step 2 — Data processing

```{r}
# -----------------------------------------------------------------------------------------
# TRAIN / VALIDATION SPLIT
# Use 75% (~117 rows) for fitting and 25% (~39 rows) for evaluation.
# We do NOT touch wheat.test until final prediction step.
# -----------------------------------------------------------------------------------------

set.seed(1234)
wheat_split    <- initial_split(wheat.train, prop = 0.75)
wheat_analysis <- training(wheat_split)   # train (75%) — used for model fitting
wheat_val      <- testing(wheat_split)    # val (25%)  — used for RMSE evaluation
```

### Step 3 — Fitting multiple candidate models (on wheat_analysis)

```{r}
set.seed(1234)
library(kknn)  

# Define all 6 model in a named list(nn1=non-standardized neural network)
models <- list(

  # Linear Discriminant Analysis
  lda = discrim_linear(),

  # Naive Bayes with kernel density estimationn,
  nb  = naive_Bayes(engine = "naivebayes"),

  # K-Nearest Neighbours (k = 5 default)
  knn = nearest_neighbor(mode = "classification", neighbors = 5),

  # Decision Tree (min_n = 5)
  rp  = decision_tree(mode = "classification", min_n = 5),

  # Random Forest
  rf  = rand_forest(mode = "classification"),
   
  #  neural network
  # p = 7 numeric predictors → ceiling(7/2) = 4 hidden units
  nn1 = mlp(mode="classification",
  hidden_units = 4, epochs = 1000))


# Fit all five models simultaneously on the analysis (training) split
fitted_models <- models |>
  map(fit, variety ~ ., data = wheat_analysis)
```

### Step 4 — Neural networks (standardise).

```{r}
set.seed(1234)
library(recipes)
# Neural networks are sensitive to scale — standardise all numeric predictors.
wheat.rec <- recipe(variety ~ ., data = wheat_analysis) |>
  step_normalize(all_numeric_predictors()) |>
  prep(wheat_analysis)

# hidden_units = ceiling(p/2) per 
# p = 7 numeric predictors → ceiling(7/2) = 4 hidden units
wheat.nn <- mlp(
  mode         = "classification",
  hidden_units = 4,      
  penalty      = 0.01,   
  epochs       = 1000    
) |>
  fit(variety ~ .,
      data = bake(wheat.rec, wheat_analysis))    

# Check for convergence
 wheat.nn |> pluck("fit", "convergence")
```

### Step 5 — Multinomial Logistic Regression.

```{r}
set.seed(1234)
wheat.mn <- multinom_reg() |>
  fit(variety ~ ., data = wheat_analysis)

wheat.mn |>
  augment(new_data = wheat_val) |>
  accuracy(truth = variety, estimate = .pred_class)

wheat.mn |>
  augment(new_data = wheat_val) |>
  conf_mat(truth = variety, estimate = .pred_class)
```

### Step 6 — Compare all models on wheat.val

```{r}
set.seed(1234)
library(recipes)
# Generate predictions for the 6 models on the validation set
val_preds <- fitted_models |>
  map(augment, new_data = wheat_val) |>
  bind_rows(.id = "model")

# Calculate validation accuracy for 6 models
val_accuracy <- val_preds |>
  group_by(model) |>
  accuracy(truth = variety, estimate = .pred_class) |>
  arrange(desc(.estimate))

# Calculate validation accuracy for the Neural Network
nn_val_preds <- wheat.nn |>
  augment(new_data = bake(wheat.rec, wheat_val))

nn_val_acc <- nn_val_preds |>
  accuracy(truth = variety, estimate = .pred_class) |>
  mutate(model = "nn")

 wheat.mn_acc <-wheat.mn |>
  augment(new_data = wheat_val) |>
  accuracy(truth = variety, estimate = .pred_class) |>
  mutate(model="Multinomial LR")

# Combine all model accuracy results to select the best model
all_val_accuracy <- bind_rows(val_accuracy, nn_val_acc,wheat.mn_acc) |>
  arrange(desc(.estimate))

print(all_val_accuracy)

# Results:
#   rp  (Decision Tree)    0.974   ← BEST
#   nb  (Naive Bayes)      0.949
#   Multinomial LR         0.949
#   rf  (Random Forest)    0.923
#   nn  (Neural Network)   0.923   ← Standardized
#   nn1                    0.897   ← No Standardized
#   lda (LDA)              0.897
#   knn (KNN k=5)          0.872

# Confusion matrices for the five non-NN models — inspect which varieties are confused
# The Decision tree model made just one incorrect prediction: one actual Kama sample was misclassified as Canadian.
fitted_models |>
  map(augment, new_data = wheat_val) |>
  map(conf_mat, truth = variety, estimate = .pred_class)

# Confusion matrix for the Neural Network
nn_val_preds |>
  conf_mat(truth = variety, estimate = .pred_class)

```

### Step 7— Refit the best model using all wheat.train rows

```{r}
library(recipes)
# Decision tree achieved the highest validation accuracy (97.4%),
# so we select it as the final mode .
best_spec <- decision_tree(mode = "classification")

# Refit on the combined training + validation set (all 156 rows of wheat.train)
final_model <- best_spec |>
  fit(variety ~ ., data = wheat.train)  

# Apply the final model to the unlabelled test set and extract predicted class labels
ex2  <- wheat.test |>
  bind_cols(
    final_model |>
      augment(new_data = wheat.test) |>
      select(.pred_class)                       
  )
# preview first 6 predictions
ex2 |> slice_head(n = 6)  

# Write submission file
ex2|>write_csv("ex2_studentid.csv")
```

### Step 8 — Accuracy evaluation after removing outliers

```{r}
set.seed(1234)
library(recipes)
#split clean data
wheat_split  <- initial_split(wheat.train.clean, prop = 0.75)
wheat_analysis_clean <- training(wheat_split )   
wheat_val_clean      <- testing(wheat_split )    

# Define all six  model (nn1:non-standardized neural network )
models_clean <- list(
  lda_clean = discrim_linear(),
  nb_clean  = naive_Bayes(engine = "naivebayes"),
  knn_clean = nearest_neighbor(mode = "classification", neighbors = 5),
  rp_clean  = decision_tree(mode = "classification", min_n = 5),
  rf_clean  = rand_forest(mode = "classification"),
  nn1_clean = mlp(mode="classification",
  hidden_units = 4, epochs = 1000))
# Multinomial Regression (
wheat.mn_clean <- multinom_reg() |>
  fit(variety ~ ., data = wheat_analysis_clean)

# Fit all six models
fitted_models_clean <- models_clean |>
  map(fit, variety ~ ., data = wheat_analysis_clean)


# Generate predictions for the 6 models on the validation set
val_preds_clean <- fitted_models_clean |>
  map(augment, new_data = wheat_val_clean) |>
  bind_rows(.id = "model")
# Calculate validation accuracy for non-NN models
val_accuracy_clean <- val_preds_clean |>
  group_by(model) |>
  accuracy(truth = variety, estimate = .pred_class) |>
  arrange(desc(.estimate))


# Neural networks— standardise
wheat.rec_clean <- recipe(variety ~ ., data = wheat_analysis_clean) |>
  step_normalize(all_numeric_predictors()) |>
  prep(wheat_analysis_clean)

wheat.nn_clean <- mlp(
  mode         = "classification",
  hidden_units = 4,      
  penalty      = 0.01,   
  epochs       = 1000    
) |>
  fit(variety ~ .,
      data = bake(wheat.rec_clean, wheat_analysis_clean))    

# Check for convergence
 wheat.nn_clean |> pluck("fit", "convergence")

# Calculate validation accuracy for the Neural Network
# Apply the same normalisation recipe to the validation set (bake uses training stats)
nn_val_preds_clean <- wheat.nn_clean |>
  augment(new_data = bake(wheat.rec_clean, wheat_val_clean))

nn_val_acc_clean <- nn_val_preds_clean |>
  accuracy(truth = variety, estimate = .pred_class) |>
  mutate(model = "nn_clean")

wheat.mn_acc_clean <-wheat.mn_clean |>
  augment(new_data = wheat_val_clean) |>
  accuracy(truth = variety, estimate = .pred_class) |>
  mutate(model="Multinomial LR_clean")

# Combine all model accuracy results to select the best performer
all_val_accuracy_clean <- bind_rows(val_accuracy_clean, nn_val_acc_clean,wheat.mn_acc_clean) |>
  arrange(desc(.estimate))

print(all_val_accuracy_clean)
# After removing outliers, the accuracy of the LDA model increased from 89.7% to 97.4%, while the accuracy of the unstandardized Neural Network model also improved from 89.7% to 97.4%. However, the accuracy of the RP model decreased from 97.4% to 84.6%. Since the outliers were not caused by input errors and similar outliers may also appear in the test set, the final model selection remains the RP model trained on the original dataset without outlier removal.
```

### 2.1 Wheat Variety Methodology

**Exploratory Analysis.** The training dataset contains `no missing`
values (1a), and each variety class has an identical count of
`52 observations` (1c). The density plots illustrate the distribution of
each predictor across the three varieties, with only kernel asymmetry,
kernel area, and groove length showing noticeable distinctions among the
classes (1e). The boxplots indicate that all variables contain
`outliers`, and these outliers are present and evenly distributed across
all three classes (1f). The pair plot reveals strong correlations among
variables, with most r > 0.97 (1g). After removing the outliers, the
resulting boxplots and density plots exhibited clearer trends (1i).

**Modelling Techniques Tried.** `Seven classification` methods were
applied using the 75% dataset for model fitting and the remaining 25%
for validation. For parameter selection, KNN used k=5, the Decision Tree
used min_n=5, and the Neural Network consisted of a hidden layer
with ⌈7/2⌉=4 units, a penalty of 0.01, and 1,000 epochs.

``Outlier removal`` was explored in Step 8, improving the neural network
accuracy (89.7%→97.4%) but reducing the Decision Tree accuracy
(97.4%→84.6%). Since outliers were not caused by data entry errors and
may also be present in the test data, the original dataset was retained.

For the ``Neural Network`` model, we compared the performance between the
standardized data and the original dataset. After standardization, the
accuracy improved notably from 89.7% to 92.3%.

Model summary:

| Model                | Validation Accuracy |
|:---------------------|:--------------------|
| Decision Tree (rp)   | 0.974               |
| Naive Bayes (nb)     | 0.949               |
| Multinomial LR       | 0.949               |
| Random Forest (rf)   | 0.923               |
| Neural Network (nn)  | 0.923               |
| LDA                  | 0.897               |
| Neural Network (nn1) | 0.897               |
| KNN (k=5)            | 0.872               |

**Final Model Selected.** Based on the accuracy results, the Decision
Tree model achieved the highest accuracy at ``97.4%``, only one
misclassification occurred (Kama predicted as Canadian). The model
achieved 100% correct separation for the ``Rosa`` and ``Canadian`` wheat
varieties.To ensure optimal performance, we trained the model(Decision
Tree) on the entire training dataset before testing.


### 2.2 Wheat Variety Explainability

#### Group means by variety(Step 1: 1h)

| Variety  | area | perimeter | compactness | length | width | asymmetry | groove |
|----------|------|-----------|-------------|--------|-------|-----------|--------|
| Canadian | 14.0 | 15.6      | 0.993       | 6.17   | 3.34  | 5.78      | 6.06   |
| Kama     | 16.9 | 16.8      | 1.03        | 6.47   | 3.81  | 3.18      | 6.00   |
| Rosa     | 21.5 | 18.9      | 1.04        | 7.16   | 4.33  | 4.13      | 7.02   |

The mean table shows Rosa has the largest kernel area (21.5), perimeter
(18.9), and groove length (7.16), while Canadian has the smallest kernel
area (14) and perimeter (15.6) but the highest asymmetry (5.78), and
Kama has the lowest asymmetry (3.18). The density plots and boxplots
(Step 1: 1i) confirm that kernel width, area, and perimeter separate all
three varieties with no overlap , and groove length
distinctly separates Rosa from the other two. Pairs Plot (Step 1: 1g)
reveals correlations exceeding 97% among most predictors, 2–3 variables
such as kernel width and groove length are sufficient for effective
classification.
